### 2021-03-04 ch/2-related-work.tex
Fix headings and remove duplicate bib entries

\chapter{Related work}\label{ch:related-work}
%
% =========================
%      old introduction
% =========================
% \section{Introduction}\label{introduction}
% Fairness in Machine Learning is a topic that has recently received increasing attention.
% Machine Learning has become very widespread in our society
% which brings some dangers with it.
% Decisions such as credit eligibility and hiring are increasingly done automatically.
% However, the algorithms that are used in these decisions can have strong biases
% that stem from the training data.
% Often, the algorithms are even more biased than the training data.
% \citet{kamishima2011fairness} give a great illustration for this
% (adapted from \citet{calders2010three}).
% The illustration is based on the Adult/Census Income dataset \citep{kohavi1996scaling}
% which contains records of individuals with thirteen attributes, among them age and gender.
% The task is to predict whether or not the individual earns more than \$50K per year.
% In the dataset, 32\% of the men have a positive label (\ie, they earn more than \$50K),
% but of the women only 11\% have a positive label.
% With this ratio of about 3:1, the dataset is already skewed.
% However, when training a \href{https://en.wikipedia.org/wiki/Naive_Bayes_classifier}{Na\"ive Bayes classifier} on the data,
% the ratio in the predictions is even more unfair:
% the amount of women who get a positive prediction is even lower than 11\%.
% If the Na\"ive Bayes classifier does not get access to the gender information,
% the predictions will be a bit fairer,
% but they are still more unfair than the original dataset labels.
%
% This phenomenon of predictions being more unfair than an already skewed dataset
% is due to the generalization that classifiers have to make.
% It is easier to classify all but the most outstanding women as low income
% than to learn a complicated decision boundary.
% Furthermore, removing the gender information does not solve the problem
% because the gender can be inferred from other data.
%
% The following is a review of the literature on Fairness in Machine Learning.
% Several different approaches to and definition of Fairness are presented.
% The focus is on identifying the major trends in the field from the beginning.
\section{\NoCaseChange{Two views of the dataset bias problem}}\label{two-views-of-the-dataset-bias-problem}
As hinted to in the introduction,
there is a divide in the relevant literature in how the problem of dataset bias is approached.
On the one hand, there is the ``ground-truth-centric'' view -- expressed through most of the introduction --
that the training set is biased but the deployment setting is not.
By \emph{deployment setting} I am here referring to the real distribution
that will be encountered by the machine learning system once it is in use.

@@

\citet{kamishima2011fairness} take a closer look
at the underlying causes for unfair predictors and condense it to three main causes:
\emph{prejudice}, \emph{underestimation} and \emph{negative legacy}.
(Let \(x\) be the non-sensitive features.)

\begin{enumerate}
\item
  Prejudice:
  a statistical dependency between the sensitive attribute
  and the class label or the (non-sensitive) input features.
  If the dependency is with the class label, then the paper calls it \emph{indirect prejudice}.
  If it is with the non-sensitive features, they call it \emph{latent prejudice}.
  In line with \citet{pedreshi2008discrimination},
  \emph{direct prejudice} refers to classifiers that make direct use of the sensitive attribute.
  Indirect prejudice and latent prejudice can be quantified by
  the mutual information between \(y\) and \(s\) or \(x\) and \(s\) respectively.
  \citet{kamishima2011fairness} do not consider latent prejudice to be immediately harmful,
  but it can be a problem with respect to the protection of personal data and compliance with laws.
\item
  Underestimation:
  non-convergence of machine learning algorithm due to limited amount of data.
  This magnitude of this problem can be estimated
  by considering the difference between the actual training sample distribution
  and the distribution that the machine learning model has internalized.
  This is the underlying cause of models that make predictions
  that are even more unfair than the dataset.
\item
  Negative Legacy:
  sampling bias and wrong labels in the dataset.
  In contrast to the problem of \emph{prejudice},
  \emph{negative legacy} might not be detectable by analysing the dataset.
  \emph{Sampling bias}, meaning that certain data points simply are missing,
  and \emph{wrong labels} can only be corrected if other sources of information are available.
\end{enumerate}

We see that previous works nearly exclusively dealt with the first of these issues, \emph{prejudice},
and more precisely \emph{indirect prejudice}.
\citet{kamishima2011fairness} also focus on \emph{indirect prejudice};
essentially because it is the easiest to deal with, apart from \emph{direct prejudice}.
They use \acf{LR} as the basis for their fair classifier.
Their approach to enforcing fairness is qualitatively different from the previous proposals.
Instead of manipulating the dataset or using explicit algorithm to make a classifier fair,
\citet{kamishima2011fairness} treat the fairness constraint like a regulariser.
A term is added to the objective function
that estimates the mutual information between \(y\) and \(s\)
and gets minimized together with the other parts of the objective function.
This strategy leaves most of the work to the optimization that would need to be run anyway.
A factor in front of the regularization term determines how much to value fairness over accuracy.

In a comparison between \citet{kamishima2011fairness}'s algorithm
and the Na\"ive Bayes method with two different models for the sensitive attributes from \citet{calders2010three},
the latter performed better on the Adult/Census Income dataset
(higher accuracy and lower discrimination score).
However, the former has the advantage
that the sensitive attribute does not necessarily have to be known for making predictions
and the trade-off between accuracy and fairness can be easily controlled
with the factor for the regularization term.

@@

The proposed idea is then as follows: a classifier is fair
if and only if the predictive distributions for any two data points
are at least as similar as the two points themselves,
according to a given similarity measure for distributions
and a given similarity measure for data points.
They call this condition the \emph{Lipschitz condition}.
The authors propose two practical similarity measures for distributions
that are well-known in the respective literature.
In order to train a fair classifier,
the Lipschitz condition is then used as a constraint for the optimization.
This is similar to how \citet{kamishima2011fairness} used the fairness criterion as a regulariser.
However, \citet{dwork2012fairness} intend to realize it as a strict constraint.
Furthermore, they show that their criterion is compatible with Demographic Parity
but does not guarantee it in general.
When the two demographic groups are similar according to the individual similarity measure,
then the Lipschitz condition will lead to a good approximation of Demographic Parity.
If the demographic groups are very different,
then the Lipschitz criterion is essentially incompatible with Demographic Parity.

@@

The experiments are done on the German credit dataset, the Adult / Census Income dataset
and a dataset based on the Heritage Health Prize milestone 1 challenge
where the goal is to predict how many days a given person will spend in the hospital in a year.
The proposed methods performs better in these experiments
than the methods from \citet{kamiran2009classifying} and the method from \citet{kamishima2011fairness}.
However, they do not compare against \citet{calders2010three},
presumably because those methods rely on the sensitive attribute for making predictions,
which can be considered an unfair advantage.

@@

In the experiments, the authors first investigate the theoretically predicted relationship
between \(\varepsilon\)-fairness and the \(DI\) measure.
All experiments were done with either of the two Partial Repair algorithms
which give very similar results.
Three datasets are used:
the German Credit dataset, the Adult/Census Income dataset
and the Ricci dataset from the \emph{Ricci vs DeStefano} court case.
The goal for the Ricci dataset is to predict which firefighters to promote based on test scores.
The sensitive attribute is \emph{race}.
The experimental results support the theoretical predictions to a reasonable extent.
To compare the method with previous work,
several classifiers are trained on the fair representation:
logistic regression, SVM and Na\"ive Bayes.
Of these, the Na\"ive Bayes classifier seems to outperform the algorithms from previous works
\citep{kamiran2009classifying,kamishima2011fairness,zemel2013learning}
in both achieved fairness and accuracy.

>>>>

\chapter{Related work}\label{ch:related-work}
%
% =========================
%      old introduction
% =========================
% \section{Introduction}\label{introduction}
% Fairness in Machine Learning is a topic that has recently received increasing attention.
% Machine Learning has become very widespread in our society
% which brings some dangers with it.
% Decisions such as credit eligibility and hiring are increasingly done automatically.
% However, the algorithms that are used in these decisions can have strong biases
% that stem from the training data.
% Often, the algorithms are even more biased than the training data.
% \citet{kamishima2012fairness} give a great illustration for this
% (adapted from \citet{calders2010three}).
% The illustration is based on the Adult/Census Income dataset \citep{kohavi1996scaling}
% which contains records of individuals with thirteen attributes, among them age and gender.
% The task is to predict whether or not the individual earns more than \$50K per year.
% In the dataset, 32\% of the men have a positive label (\ie, they earn more than \$50K),
% but of the women only 11\% have a positive label.
% With this ratio of about 3:1, the dataset is already skewed.
% However, when training a \href{https://en.wikipedia.org/wiki/Naive_Bayes_classifier}{Na\"ive Bayes classifier} on the data,
% the ratio in the predictions is even more unfair:
% the amount of women who get a positive prediction is even lower than 11\%.
% If the Na\"ive Bayes classifier does not get access to the gender information,
% the predictions will be a bit fairer,
% but they are still more unfair than the original dataset labels.
%
% This phenomenon of predictions being more unfair than an already skewed dataset
% is due to the generalization that classifiers have to make.
% It is easier to classify all but the most outstanding women as low income
% than to learn a complicated decision boundary.
% Furthermore, removing the gender information does not solve the problem
% because the gender can be inferred from other data.
%
% The following is a review of the literature on Fairness in Machine Learning.
% Several different approaches to and definition of Fairness are presented.
% The focus is on identifying the major trends in the field from the beginning.
\section{\NoCaseChange{Two views of the dataset bias problem}}\label{two-views-of-the-dataset-bias-problem}
As hinted to in the introduction,
there is a divide in the relevant literature in how the problem of dataset bias is approached.
On the one hand, there is the ``ground-truth-centric'' view -- expressed through most of the introduction --
that the training set is biased but the deployment setting is not.
By \emph{deployment setting} I am here referring to the real distribution
that will be encountered by the machine learning system once it is in use.

@@

\citet{kamishima2012fairness} take a closer look
at the underlying causes for unfair predictors and condense it to three main causes:
\emph{prejudice}, \emph{underestimation} and \emph{negative legacy}.
(Let \(x\) be the non-sensitive features.)

\begin{enumerate}
\item
  Prejudice:
  a statistical dependency between the sensitive attribute
  and the class label or the (non-sensitive) input features.
  If the dependency is with the class label, then the paper calls it \emph{indirect prejudice}.
  If it is with the non-sensitive features, they call it \emph{latent prejudice}.
  In line with \citet{pedreshi2008discrimination},
  \emph{direct prejudice} refers to classifiers that make direct use of the sensitive attribute.
  Indirect prejudice and latent prejudice can be quantified by
  the mutual information between \(y\) and \(s\) or \(x\) and \(s\) respectively.
  \citet{kamishima2012fairness} do not consider latent prejudice to be immediately harmful,
  but it can be a problem with respect to the protection of personal data and compliance with laws.
\item
  Underestimation:
  non-convergence of machine learning algorithm due to limited amount of data.
  This magnitude of this problem can be estimated
  by considering the difference between the actual training sample distribution
  and the distribution that the machine learning model has internalized.
  This is the underlying cause of models that make predictions
  that are even more unfair than the dataset.
\item
  Negative Legacy:
  sampling bias and wrong labels in the dataset.
  In contrast to the problem of \emph{prejudice},
  \emph{negative legacy} might not be detectable by analysing the dataset.
  \emph{Sampling bias}, meaning that certain data points simply are missing,
  and \emph{wrong labels} can only be corrected if other sources of information are available.
\end{enumerate}

We see that previous works nearly exclusively dealt with the first of these issues, \emph{prejudice},
and more precisely \emph{indirect prejudice}.
\citet{kamishima2012fairness} also focus on \emph{indirect prejudice};
essentially because it is the easiest to deal with, apart from \emph{direct prejudice}.
They use \acf{LR} as the basis for their fair classifier.
Their approach to enforcing fairness is qualitatively different from the previous proposals.
Instead of manipulating the dataset or using explicit algorithm to make a classifier fair,
\citet{kamishima2012fairness} treat the fairness constraint like a regulariser.
A term is added to the objective function
that estimates the mutual information between \(y\) and \(s\)
and gets minimized together with the other parts of the objective function.
This strategy leaves most of the work to the optimization that would need to be run anyway.
A factor in front of the regularization term determines how much to value fairness over accuracy.

In a comparison between \citet{kamishima2012fairness}'s algorithm
and the Na\"ive Bayes method with two different models for the sensitive attributes from \citet{calders2010three},
the latter performed better on the Adult/Census Income dataset
(higher accuracy and lower discrimination score).
However, the former has the advantage
that the sensitive attribute does not necessarily have to be known for making predictions
and the trade-off between accuracy and fairness can be easily controlled
with the factor for the regularization term.

@@

The proposed idea is then as follows: a classifier is fair
if and only if the predictive distributions for any two data points
are at least as similar as the two points themselves,
according to a given similarity measure for distributions
and a given similarity measure for data points.
They call this condition the \emph{Lipschitz condition}.
The authors propose two practical similarity measures for distributions
that are well-known in the respective literature.
In order to train a fair classifier,
the Lipschitz condition is then used as a constraint for the optimization.
This is similar to how \citet{kamishima2012fairness} used the fairness criterion as a regulariser.
However, \citet{dwork2012fairness} intend to realize it as a strict constraint.
Furthermore, they show that their criterion is compatible with Demographic Parity
but does not guarantee it in general.
When the two demographic groups are similar according to the individual similarity measure,
then the Lipschitz condition will lead to a good approximation of Demographic Parity.
If the demographic groups are very different,
then the Lipschitz criterion is essentially incompatible with Demographic Parity.

@@

The experiments are done on the German credit dataset, the Adult / Census Income dataset
and a dataset based on the Heritage Health Prize milestone 1 challenge
where the goal is to predict how many days a given person will spend in the hospital in a year.
The proposed methods performs better in these experiments
than the methods from \citet{kamiran2009classifying} and the method from \citet{kamishima2012fairness}.
However, they do not compare against \citet{calders2010three},
presumably because those methods rely on the sensitive attribute for making predictions,
which can be considered an unfair advantage.

@@

In the experiments, the authors first investigate the theoretically predicted relationship
between \(\varepsilon\)-fairness and the \(DI\) measure.
All experiments were done with either of the two Partial Repair algorithms
which give very similar results.
Three datasets are used:
the German Credit dataset, the Adult/Census Income dataset
and the Ricci dataset from the \emph{Ricci vs DeStefano} court case.
The goal for the Ricci dataset is to predict which firefighters to promote based on test scores.
The sensitive attribute is \emph{race}.
The experimental results support the theoretical predictions to a reasonable extent.
To compare the method with previous work,
several classifiers are trained on the fair representation:
logistic regression, SVM and Na\"ive Bayes.
Of these, the Na\"ive Bayes classifier seems to outperform the algorithms from previous works
\citep{kamiran2009classifying,kamishima2012fairness,zemel2013learning}
in both achieved fairness and accuracy.

### 2021-03-04 paper1/paper1.tex
Fix headings and remove duplicate bib entries

% \newcommand{P}{\mathbb{P}} % Probability
% \renewcommand{\algorithmicrequire}{\textbf{Input:}}
% \renewcommand{\algorithmicensure}{\textbf{Output:}}
% \newcommand{\thomas}[1]{\textcolor{brown}{\textbf{Thomas:} #1}}
% \newcommand{\novi}[1]{\textcolor{red}{\textbf{Novi:} #1}}
% \newcommand{\myles}[1]{\textcolor{blue}{\textbf{Myles:} #1}}

@@

There are many existing models for enforcing demographic parity and equalised odds
\citep{CreMadJacWeietal19,AgaBeyDudLanetal18,calders2009building,kamishima2012fairness,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.
However, these existing approaches to balancing accuracy and fairness rely on intermediate, unintuitive control parameters such as allowable constraint violation $\epsilon$ (e.g. $0.01$) in \citet{AgaBeyDudLanetal18}, or a covariance threshold $c$
(e.g. $0$ that is controlled by another parameters $\tau$ and $\mu$ -- $0.005$ and $1.2$ -- to trade off this threshold and accuracy)
in \citet{zafar2017fairnesstreatment}.
This is related to the fact that many of these approaches embed fairness criteria as \emph{constraints} in the optimisation procedure
\citep{DonOneBenShaetal18,quadrianto2017recycling,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.

@@

\subsubsection{Choosing a target rate.}
\noindent As shown, there is a remaining degree of freedom when targeting a balanced dataset:
the target rate $\mathit{PR}_t := P(\bar{y}=1)$.
This is true for both fairness criteria that we are targeting.
The choice of targeting rate affects how much $\eta$ and $\bar{\eta}$ differ as implied by Theorem~\ref{th:prob}
($\mathit{PR}_t$ affects $m_s$ and $b_s$).
$\bar{\eta}$ should remain close to $\eta$
as $\bar{\eta}$ only represents an auxiliary distribution that does not have meaning on its own.
The threshold $t_s$ in Theorem~\ref{th:prob} (\eqref{eq:def-ts}) gives an indication of how close the distributions are.
With the definitions in \eqref{eq:tau-11} and \eqref{eq:tau-01},
we can express $t_s$ in terms of the target rate and the base rate:
\begin{align}
  t_s = \begin{cases}
    \frac{1}{2}\frac{\mathit{PR}_b^s - \mathit{PR}_t}{\mathit{PR}_t} &\text{if }\mathit{PR}_t>\mathit{PR}_b^j\\
    \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^s}{1 - \mathit{PR}_t} &\text{otherwise.}
  \end{cases}\label{eq:ts-pr}
\end{align}
This shows that $t_s$ is smallest when $\mathit{PR}_b^s$ and $\mathit{PR}_t$ are closest.
However, as $\mathit{PR}_b^s$ has different values for different $s$,
we cannot set $\mathit{PR}_b^s=\mathit{PR}_t$ for all $s$.
In order to keep both $t_{s=0}$ and $t_{s=1}$ small,
it follows from \eqref{eq:ts-pr} that $\mathit{PR}_t$ should at least be between $\mathit{PR}_b^0$ and $\mathit{PR}_b^1$.
A more precise statement can be made when we explicitly want to minimise the sum $t_{s=0} + t_{s=1}$:
assuming $\mathit{PR}_b^0<\mathit{PR}_t<\mathit{PR}_b^1$ and $\mathit{PR}_b^1<\tfrac{1}{2}$,
the optimal choice for $\mathit{PR}_t$ is $\mathit{PR}_b^1$ (see Supplementary Material for details).
We call this choice $\mathit{PR}_t^{max}$.
For $\mathit{PR}_b^0>\tfrac{1}{2}$, analogous statements can be made,
but this is of less interest as this case does not appear in our experiments.

@@

\subsection{Conditionally balanced dataset}
\noindent There is a fairness definition related to demographic parity which allows conditioning on ``legitimate'' risk factors $\ell$
when considering how equal the demographic groups are treated~\citep{corbett2017algorithmic}.
This cleanly translates into balanced datasets which are balanced conditioned on $\ell$:
\begin{align}
  &P(\bar{y}=1|\ell=\ell\prime, s=0) \overset{!}{=} P(\bar{y}=1|\ell=\ell\prime, s=1)~.
\end{align}
We can interpret this as splitting the data into partitions based on the value of $\ell$,
where the goal is to have all these partitions be balanced.
This can easily be achieved by our method by setting a $\mathit{PR}_t(\ell)$ for each value of $\ell$
and computing the transition probabilities for each sample depending on $\ell$.

\section{Related work}\label{sec:relatedwork}
There are several ways to enforce fairness in machine learning models:
as a pre-processing step
% (e.g.~),
\citep{kamiran2012data,louizos2016variational,lum2016statistical,zemel2013learning,Chiappa19,QuaShaTho19},
% kusner2017counterfactual
as a post-processing step \citep{feldman2015certifying,hardt2016equality}, %(e.g.~\citep{feldman2015certifying,hardt2016equality}),
 or as a constraint during the learning phase
% (e.g.~).
\citep{calders2009building,zafar2017fairnesstreatment,zafar2017fairnessconstraints,DonOneBenShaetal18,DimLiuParRad19}.
%
Our method enforces fairness during the learning phase (an in-processing approach) but, unlike other approaches, we do not cast fair-learning as a \emph{constrained} optimisation problem.
Constrained optimisation requires a customised procedure.
%
In \citet{GohCotGupFri16}, \citet{zafar2017fairnessconstraints}, and \citet{zafar2017fairnesstreatment},
% In~\citet{GohCotGupFri16,zafar2017fairnesstreatment,zafar2017fairnessconstraints},
suitable majorisation-minimisation/convex-concave procedures~\citep{LanSri09} were derived.
%
Furthermore, such constrained optimisation approaches may lead to more unstable training,
and often yield classifiers with both worse accuracy and more unfair~\citep{CotJiaWanNar18}.

@@

The fair GP models and the baseline GP model are all based on variational inference and use the same settings.
During training, each batch is equivalent to the whole dataset.
The number of inducing inputs is 500 on the ProPublica dataset
and 2500 on the Adult dataset
which corresponds to approximately $\nicefrac{1}{8}$ of the number of training points for each dataset.
We use a squared-exponential (SE) kernel with automatic relevance determination (ARD)
and the probit function as the likelihood function.
We optimise the hyper-parameters and the variational parameters
using the Adam method~\citep{kingma2014adam} with the default parameters. %($\beta_1 = 0.9$, $\beta_2 = 0.999$).
We use the full covariance matrix for the Gaussian variational distribution.
%
The logistic regression is trained with RAdam~\citep{liu2019variance} and uses L2 regularisation.
% The regularization coefficient differs between datasets but not between repeats.
For the regularisation coefficient, we conducted a hyper-parameter search over 10 folds of the data.
For each fold, we picked the hyper-parameter which achieved the best fairness among those 5 with
the best accuracy scores.
We then averaged over the 10 hyper-parameter values chosen in this way and then used this average for all runs to obtain our final results.
% For the Adult dataset, the regularization parameter is $0.0024$ for ProPublica and $0.00035$ for Adult.

>>>>

% \newcommand{P}{\mathbb{P}} % Probability

@@

There are many existing models for enforcing demographic parity and equalised odds
\citep{creager2019flexibly,AgaBeyDudLanetal18,calders2009building,kamishima2012fairness,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.
However, these existing approaches to balancing accuracy and fairness rely on intermediate, unintuitive control parameters such as allowable constraint violation $\epsilon$ (e.g. $0.01$) in \citet{AgaBeyDudLanetal18}, or a covariance threshold $c$
(e.g. $0$ that is controlled by another parameters $\tau$ and $\mu$ -- $0.005$ and $1.2$ -- to trade off this threshold and accuracy)
in \citet{zafar2017fairnesstreatment}.
This is related to the fact that many of these approaches embed fairness criteria as \emph{constraints} in the optimisation procedure
\citep{DonOneBenShaetal18,quadrianto2017recycling,zafar2017fairnesstreatment,zafar2017fairnessconstraints}.

@@

\paragraph{Choosing a target rate.}
As shown, there is a remaining degree of freedom when targeting a balanced dataset:
the target rate $\mathit{PR}_t := P(\bar{y}=1)$.
This is true for both fairness criteria that we are targeting.
The choice of targeting rate affects how much $\eta$ and $\bar{\eta}$ differ as implied by Theorem~\ref{th:prob}
($\mathit{PR}_t$ affects $m_s$ and $b_s$).
$\bar{\eta}$ should remain close to $\eta$
as $\bar{\eta}$ only represents an auxiliary distribution that does not have meaning on its own.
The threshold $t_s$ in Theorem~\ref{th:prob} (\eqref{eq:def-ts}) gives an indication of how close the distributions are.
With the definitions in \eqref{eq:tau-11} and \eqref{eq:tau-01},
we can express $t_s$ in terms of the target rate and the base rate:
\begin{align}
  t_s = \begin{cases}
    \frac{1}{2}\frac{\mathit{PR}_b^s - \mathit{PR}_t}{\mathit{PR}_t} &\text{if }\mathit{PR}_t>\mathit{PR}_b^j\\
    \frac{1}{2}\frac{\mathit{PR}_t - \mathit{PR}_b^s}{1 - \mathit{PR}_t} &\text{otherwise.}
  \end{cases}\label{eq:ts-pr}
\end{align}
This shows that $t_s$ is smallest when $\mathit{PR}_b^s$ and $\mathit{PR}_t$ are closest.
However, as $\mathit{PR}_b^s$ has different values for different $s$,
we cannot set $\mathit{PR}_b^s=\mathit{PR}_t$ for all $s$.
In order to keep both $t_{s=0}$ and $t_{s=1}$ small,
it follows from \eqref{eq:ts-pr} that $\mathit{PR}_t$ should at least be between $\mathit{PR}_b^0$ and $\mathit{PR}_b^1$.
A more precise statement can be made when we explicitly want to minimise the sum $t_{s=0} + t_{s=1}$:
assuming $\mathit{PR}_b^0<\mathit{PR}_t<\mathit{PR}_b^1$ and $\mathit{PR}_b^1<\tfrac{1}{2}$,
the optimal choice for $\mathit{PR}_t$ is $\mathit{PR}_b^1$ (see Supplementary Material for details).
We call this choice $\mathit{PR}_t^{max}$.
For $\mathit{PR}_b^0>\tfrac{1}{2}$, analogous statements can be made,
but this is of less interest as this case does not appear in our experiments.

@@

\subsection{Conditionally balanced dataset}
There is a fairness definition related to demographic parity which allows conditioning on ``legitimate'' risk factors $\ell$
when considering how equal the demographic groups are treated~\citep{corbett2017algorithmic}.
This cleanly translates into balanced datasets which are balanced conditioned on $\ell$:
\begin{align}
  &P(\bar{y}=1|\ell=\ell\prime, s=0) \overset{!}{=} P(\bar{y}=1|\ell=\ell\prime, s=1)~.
\end{align}
We can interpret this as splitting the data into partitions based on the value of $\ell$,
where the goal is to have all these partitions be balanced.
This can easily be achieved by our method by setting a $\mathit{PR}_t(\ell)$ for each value of $\ell$
and computing the transition probabilities for each sample depending on $\ell$.

\section{Related work}\label{sec:relatedwork}
There are several ways to enforce fairness in machine learning models:
as a pre-processing step
% (e.g.~),
\citep{kamiran2012data,louizos2016variational,lum2016statistical,zemel2013learning,chiappa2019path,QuaShaTho19},
% kusner2017counterfactual
as a post-processing step \citep{feldman2015certifying,hardt2016equality}, %(e.g.~\citep{feldman2015certifying,hardt2016equality}),
 or as a constraint during the learning phase
% (e.g.~).
\citep{calders2009building,zafar2017fairnesstreatment,zafar2017fairnessconstraints,DonOneBenShaetal18,DimLiuParRad19}.
%
Our method enforces fairness during the learning phase (an in-processing approach) but, unlike other approaches, we do not cast fair-learning as a \emph{constrained} optimisation problem.
Constrained optimisation requires a customised procedure.
%
In \citet{GohCotGupFri16}, \citet{zafar2017fairnessconstraints}, and \citet{zafar2017fairnesstreatment},
% In~\citet{GohCotGupFri16,zafar2017fairnesstreatment,zafar2017fairnessconstraints},
suitable majorisation-minimisation/convex-concave procedures~\citep{LanSri09} were derived.
%
Furthermore, such constrained optimisation approaches may lead to more unstable training,
and often yield classifiers with both worse accuracy and more unfair~\citep{CotJiaWanNar18}.

@@

The fair GP models and the baseline GP model are all based on variational inference and use the same settings.
During training, each batch is equivalent to the whole dataset.
The number of inducing inputs is 500 on the ProPublica dataset
and 2500 on the Adult dataset
which corresponds to approximately $\nicefrac{1}{8}$ of the number of training points for each dataset.
We use a squared-exponential (SE) kernel with automatic relevance determination (ARD)
and the probit function as the likelihood function.
We optimise the hyper-parameters and the variational parameters
using the Adam method~\citep{kingma2015adam} with the default parameters. %($\beta_1 = 0.9$, $\beta_2 = 0.999$).
We use the full covariance matrix for the Gaussian variational distribution.
%
The logistic regression is trained with RAdam~\citep{liu2019variance} and uses L2 regularisation.
% The regularization coefficient differs between datasets but not between repeats.
For the regularisation coefficient, we conducted a hyper-parameter search over 10 folds of the data.
For each fold, we picked the hyper-parameter which achieved the best fairness among those 5 with
the best accuracy scores.
We then averaged over the 10 hyper-parameter values chosen in this way and then used this average for all runs to obtain our final results.
% For the Adult dataset, the regularization parameter is $0.0024$ for ProPublica and $0.00035$ for Adult.

### 2021-03-04 paper2/1-introduction.tex
Fix headings and remove duplicate bib entries

In the typical setup, we have an input $\bm{x}$, a sensitive attribute $s$ that represents some non-admissible information like gender
and a class label $y$ which is the prediction target.
The idea of fair \emph{representation} learning \citep{ZemWuSwePitetal13,edwards2016censoring,madras2018learning}
is then to transform the input $\bm{x}$ to a representation $\bm{z}$ which is invariant to $s$.
Thus, learning from $\bm{z}$ will not introduce a forbidden dependence on $s$.
% $\bm{z}$ can then be used to learn a predictor
% In fair representation learning \cite{ZemWuSwePitetal13}\cite{edwards2016censoring}\cite{madras2018learning}, the goal is then to find a representation $\bm{z}$ that is invariant to $s$.
A good fair representation is one that preserves most of the information from $\bm{x}$ while satisfying the aforementioned constraints.
% While fair \emph{representations} are by no means the only method of achieving this goal, approaches that explicitly depend on the target label \cite{kamiran2012data} are restrictive in that they do not readily admit transfer to unseen tasks, and are bound by the requirement of having $s$ and $y$ labels be present together.

@@

Our contribution is thus two-fold:
1) We propose a simple approach to generating representations 
% through use of an INN, 
that are invariant to a feature $s$, while having the benefit of interpretability that comes with being in the data domain.
We call our model \emph{NIFR} (\textbf{N}ull-sampling for \textbf{I}nterpretable and \textbf{F}air \textbf{R}epresentations).
2) We explore a setting where the labelled training set suffers from varying levels of sampling bias,
% which we expect to be common not only in fairness problems but machine learning problems more broadly,
demonstrating an approach based on transferring information from a more diverse representative set,
with guarantees of the non-spurious information being preserved.

>>>>

In the typical setup, we have an input $\bm{x}$, a sensitive attribute $s$ that represents some non-admissible information like gender
and a class label $y$ which is the prediction target.
The idea of fair \emph{representation} learning \citep{zemel2013learning,edwards2016censoring,madras2018learning}
is then to transform the input $\bm{x}$ to a representation $\bm{z}$ which is invariant to $s$.
Thus, learning from $\bm{z}$ will not introduce a forbidden dependence on $s$.
% $\bm{z}$ can then be used to learn a predictor
% In fair representation learning \cite{zemel2013learning}\cite{edwards2016censoring}\cite{madras2018learning}, the goal is then to find a representation $\bm{z}$ that is invariant to $s$.
A good fair representation is one that preserves most of the information from $\bm{x}$ while satisfying the aforementioned constraints.
% While fair \emph{representations} are by no means the only method of achieving this goal, approaches that explicitly depend on the target label \cite{kamiran2012data} are restrictive in that they do not readily admit transfer to unseen tasks, and are bound by the requirement of having $s$ and $y$ labels be present together.

@@

Our contribution is thus two-fold:
1) We propose a simple approach to generating representations 
% through use of an INN, 
that are invariant to a feature $s$, while having the benefit of interpretability that comes with being in the data domain.
We call our model \emph{NIFR} (\emph{N}ull-sampling for \emph{I}nterpretable and \emph{F}air \emph{R}epresentations).
2) We explore a setting where the labelled training set suffers from varying levels of sampling bias,
% which we expect to be common not only in fairness problems but machine learning problems more broadly,
demonstrating an approach based on transferring information from a more diverse representative set,
with guarantees of the non-spurious information being preserved.
